---
title: "p8105_hw2_xs2330"
author: "Xiao Shi"
date: "September 26, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

## Problem 1
*Import, select, and tidy data*
```{r prob1 import data, message=FALSE}
NYC_Transit = 
  read_csv("./data_hw2/nyc_subway.csv") %>%
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude, route1, route2, route3, route4, route5, 
         route6, route7, route8, route9, route10, route11, entry, vending, entrance_type, ada)
NYC_Transit$entry = ifelse(NYC_Transit$entry == 'YES', TRUE,FALSE)
```

The `NYC_transit` dataset contains detailed information on station, routing, staffing, ada compliance, and vending. I used the janitor package to clean up the original data, and used the select function to choose the variables that will be focused on for this analysis. The dimension is `r nrow(NYC_Transit)` rows by `r ncol(NYC_Transit)` columns. However, the data is NOT tidy yet because the route number (variables route1 to route11) is spread across 11 columns.
